Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-04-30 concurrency/CONCUR 2023/paper.tex
reformatted give-up proof
finished annotated of insert functions for both template and add them in appendix

\begin{mathpar}
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{insert(p, k, v)}\ 
	{\color{blue}
		\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \texttt{v}])
		\right\rangle
	}
\\
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{lookup(p, k)}\ 
	{\color{blue}
		\left\langle \texttt{v}. \ \
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m \land m(\texttt{k}) = \texttt{v}
		\right\rangle
	}
\\
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{delete(p, k)}\ 
	{\color{blue}
		\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \_ ])
		\right\rangle
	}
\end{mathpar}
The intention of the template approach is that we should be able to derive implementations from these operations from \lstinline{traverse} instantiated with an appropriate \lstinline{findNext} function, and prove them correct using the \lstinline{traverse} specs of the previous section, without any further reasoning about specific concurrency patterns.

@@

\begin{figure}[h]
	\begin{subfigure}{\textwidth}
	\centering
	\begin{mathpar}
		{\color{blue}
			\left\{ 
			\begin{array}{c}
				\ \texttt{pn->n->t}  
			\end{array}
			\right\}
		}
		\\ 
		\texttt{insertOp(pn, k, v)} 
		\\
		{\color{blue}
			\left\{\exists \  \texttt{t' l r lk1 lk2}.
			\begin{array}{c}
				\ \texttt{pn->n->t'} \ \ast \ \texttt{t'}\mapsto \texttt{(k, v, l, r)} \ \ast \ \\ 
				\ \texttt{l} \mapsto \texttt{(NULL, lk1)}\  \ast \ \texttt{r} \mapsto \texttt{(NULL, lk2)}
			\end{array}
			\right\}
		}
		\end{mathpar}
		\caption{Specification of \texttt{insertOp} for the lock-coupling template}
			\label{fig:insertOp_lock}
	\end{subfigure}
	%insertOp for give-up of give-up
	\begin{subfigure}{\textwidth}
		\begin{mathpar}
	{\color{blue}
		\left\{ 
		\begin{array}{c}
			\texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{t} \  \ast \ \texttt{n}\mapsto\texttt{min} \ \ast \ \texttt{n}\mapsto\texttt{max}
		\end{array}
		\right\}
	}
	\\
	\texttt{insertOp(pn, k, v)}
	\\
	{\color{blue}
		\left\{\exists \ \texttt{t' l r lk1 lk2}. 
		\begin{array}{c}
			\ \texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{t'} \ \ast \ \texttt{t'}\mapsto \texttt{(k, v, l, r)} \\ 
			\ \ast \ \texttt{l} \mapsto \texttt{(NULL, lk1, (min, k))} \ \ast \ \texttt{r} \mapsto \texttt{(NULL, lk2, (k, max))}
		\end{array}
		\right\}
	}
	\end{mathpar}
	\caption{Specification of \texttt{insertOp} for the give-up template}
	\label{fig:insertOp_giveup}
\end{subfigure}
	\caption{Specification of \texttt{insertOp} for the lock-coupling template}
	\label{fig:insertOp}
\end{figure}

@@

The atomic specification for \texttt{insert} method is
\begin{mathpar}
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{insert(p, k, v)}\ 
	{\color{blue}
		\left\langle 
		\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \texttt{v}])
		\right\rangle
	}
\end{mathpar}
\wm{Consider skipping this paragraph if we need space.} Similar to the \texttt{traverse} specification, the \texttt{insert} function's specification is composed of two components: local and public pre- and postconditions. The local pre- and postconditions, containing $\nodeboxrep \ \texttt{p}$, describe the binary search tree's concrete state with the root node $\texttt{p}$, which points to the node's lock and lock invariant assertion. On the other hand, the public pre- and postconditions are represented by the predicate $\treerep \ m$, illustrating the binary search tree's abstract state as a collection of ghost states, with each state providing information about an individual node in the tree. Here, $m$ and $m[\texttt{k} \mapsto \texttt{v}]$ denote the tree's abstract states before and after the function's execution, respectively.

The proof for \texttt{insert} begins with the precondition $\nodeboxrep \ \texttt{p}$. We set \texttt{pn->n} to point to the root node \texttt{*p} (line 3 in Figure \ref{insert_lock}). After the lock for the root node is acquired (line 4 in Figure \ref{insert_lock}), the thread acquires the contents of the root node, $\mathsf{R}(\texttt{pn->n})$. We can then satisfy the precondition of the lock-coupling \lstinline{traverse} function from Section \ref{traverse_proof_lock} \wm{Maybe we should show the spec again here}, allowing us to call \lstinline{traverse} and obtain its postcondition.

@@

\begin{figure}[h]
	\centering
	\begin{mathpar}
		% inRange for give-up
		{\color{blue}
			\left\{ 
			\begin{array}{c}
				\texttt{pn->n}\mapsto\texttt{(t, lock, min, max)} 
			\end{array}
			\right\}
		}
		\\ 
		\texttt{inRange(pn, k)} 
		\\
		{\color{blue}
			\left\{\mathit{res.} \
			\begin{array}{c}
				\ \texttt{pn->n}\mapsto\texttt{(t, lock, min, max)} \  \ast  \ 
				\\(\mathsf{if}\ \mathit{res}\ \mathsf{then}\ (\texttt{min} < \texttt{k} < \texttt{max})\ \mathsf{else}\ (\texttt{k} \leq \texttt{min} \lor \texttt{k} \geq \texttt{max}))
			\end{array}
			\right\}
		}
	\end{mathpar}
	\caption{Specifications of \texttt{inRange} and \texttt{insertOp} for the give-up template}
	\label{fig:inRange_giveup}
\end{figure}

@@

\begin{table}[h]
	\begin{center}
		\begin{tabular}{l | c || l | c } 
			\hline
			\textbf{Lock-coupling template} & \textbf{Line of proofs} & \textbf{Give-up  template} & \textbf{Line of proofs} \\
			\hline
			Template specfications & 190 & Template specfications & 138 \\
			\texttt{findNext} proof 	& 27  & \texttt{findNext} proof & 27 \\
			\texttt{insertOp} proof 	& 32  & \texttt{insertOp} proof & 31 \\
			& & \texttt{inRange} proof & 30 \\
			\texttt{traverse} proof	& 360  & \texttt{traverse} proof & 336 \\ 
			\texttt{insert} proof 	   		& 450  & \texttt{insert} proof & 284 \\
			\texttt{lookup} proof 	   	& 235  &  \texttt{lookup} proof & 304 \\ 
			\texttt{pushdown\_left} proof 	   	& 282 & & \\	
			\texttt{delete} proof 		& 182	 &   & 	 \\
			supporting proofs    & 	1909	 & supporting proofs 	& 1876 \\
			\hline
			\textbf{Total} & \textbf{3667} & & \textbf{3026}  
		\end{tabular}
		\caption{Size of Coq definitions and proofs, by topic.}
		\label{table_coq}
	\end{center}
\end{table}

@@

\section{Give-up Proof}
\label{sec:apd_giveup_proof}

@@

				(\mathit{res} = \mathsf{true} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->t} = \texttt{NULL}  \ast \infp (\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \texttt{k} \in \mathsf{range}(\texttt{p'}) \ast \cdots)

@@

%List of different predefined enumeration styles:

>>>>

\begin{mathpar}
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{insert(r, k, v)}\ 
	{\color{blue}
		\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \texttt{v}])
		\right\rangle
	}
\\
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{lookup(r, k)}\ 
	{\color{blue}
		\left\langle \texttt{v}. \ \
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ m \land m(\texttt{k}) = \texttt{v}
		\right\rangle
	}
\\
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{delete(r, k)}\ 
	{\color{blue}
		\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \_ ])
		\right\rangle
	}
\end{mathpar}
The intention of the template approach is that we should be able to derive implementations from these operations from \lstinline{traverse} instantiated with an appropriate \lstinline{findNext} function, and prove them correct using the \lstinline{traverse} specs of the previous section, without any further reasoning about specific concurrency patterns.

@@

\begin{figure}[h]
	\begin{subfigure}{\textwidth}
	\centering
	\begin{mathpar}
		{\color{blue}
			\left\{ 
			\begin{array}{c}
				\ \texttt{pn->n} \mapsto \texttt{NULL}  
			\end{array}
			\right\}
		}
		\\ 
		\texttt{insertOp(pn, k, v)} 
		\\
		{\color{blue}
			\left\{\exists \  \texttt{t l r lk1 lk2}.
			\begin{array}{c}
				\ \texttt{pn->n->t} \mapsto \texttt{(k, v, l, r)} \ \ast \ \\ 
				\ \texttt{l} \mapsto \texttt{(NULL, lk1)}\  \ast \ \texttt{r} \mapsto \texttt{(NULL, lk2)}
			\end{array}
			\right\}
		}
		\end{mathpar}
		\caption{Specification of \texttt{insertOp} for the lock-coupling template}
			\label{fig:insertOp_lock}
	\end{subfigure}
	%insertOp for give-up of give-up
	\begin{subfigure}{\textwidth}
		\begin{mathpar}
	{\color{blue}
		\left\{ 
		\begin{array}{c}
			\texttt{pn->n}\mapsto \texttt{NULL} \  \ast \ \texttt{pn->n}\mapsto \texttt{min} \ \ast \ \texttt{pn->n}\mapsto\texttt{max}
		\end{array}
		\right\}
	}
	\\
	\texttt{insertOp(pn, k, v)}
	\\
	{\color{blue}
		\left\{\exists \ \texttt{t l r lk1 lk2}. 
		\begin{array}{c}
			\texttt{pn->n->t}\mapsto \texttt{(k, v, l, r)} \  \ast \ \texttt{pn->n}\mapsto \texttt{min} \ \ast \ \texttt{pn->n}\mapsto\texttt{max}  \ \ast \ 
			\\ \texttt{l} \mapsto \texttt{(NULL, lk1, (min, k))} \ \ast \ \texttt{r} \mapsto \texttt{(NULL, lk2, (k, max))}
		\end{array}
		\right\}
	}
	\end{mathpar}
	\caption{Specification of \texttt{insertOp} for the give-up template}
	\label{fig:insertOp_giveup}
\end{subfigure}
	\caption{Specification of \texttt{insertOp} for the lock-coupling and give-up templates}
	\label{fig:insertOp}
\end{figure}

@@

The atomic specification for \texttt{insert} method is
\begin{mathpar}
	{\color{blue}
		m.\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ m
		\right\rangle
	}
	\ \texttt{insert(r, k, v)}\ 
	{\color{blue}
		\left\langle 
		\nodeboxrep\ \texttt{r} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \texttt{v}])
		\right\rangle
	}
\end{mathpar}
\wm{Consider skipping this paragraph if we need space.} Similar to the \texttt{traverse} specification, the \texttt{insert} function's specification is composed of two components: local and public pre- and postconditions. The local pre- and postconditions, containing $\nodeboxrep \ \texttt{r}$, describe the binary search tree's concrete state with the root node $\texttt{r}$, which points to the node's lock and lock invariant assertion. On the other hand, the public pre- and postconditions are represented by the predicate $\treerep \ m$, illustrating the binary search tree's abstract state as a collection of ghost states, with each state providing information about an individual node in the tree. Here, $m$ and $m[\texttt{k} \mapsto \texttt{v}]$ denote the tree's abstract states before and after the function's execution, respectively.

The proof for \texttt{insert} begins with the precondition $\nodeboxrep \ \texttt{p}$. We set \texttt{pn->n} to point to the root node \texttt{*r} (line 3 in Figure \ref{insert_lock}). After the lock for the root node is acquired (line 4 in Figure \ref{insert_lock}), the thread acquires the contents of the root node, $\mathsf{R}(\texttt{r})$. We can then satisfy the precondition of the lock-coupling \lstinline{traverse} function from Section \ref{traverse_proof_lock} \wm{Maybe we should show the spec again here}, allowing us to call \lstinline{traverse} and obtain its postcondition.

@@

\begin{figure}[h]
	\centering
	\begin{mathpar}
		% inRange for give-up
		{\color{blue}
			\left\{ 
			\begin{array}{c}
				\texttt{pn->n}\mapsto\texttt{t} \ \ast \ \texttt{pn->n}\mapsto\texttt{min} \ \ast \  \texttt{pn->n}\mapsto\texttt{max} 
			\end{array}
			\right\}
		}
		\\ 
		\texttt{inRange(pn, k)} 
		\\
		{\color{blue}
			\left\{\mathit{res.} \
			\begin{array}{c}
				\texttt{pn->n}\mapsto\texttt{t} \ \ast \ \texttt{pn->n}\mapsto\texttt{min} \ \ast \  \texttt{pn->n}\mapsto\texttt{max} \  \ast  \ 
				\\(\mathsf{if}\ \mathit{res}\ \mathsf{then}\ (\texttt{min} < \texttt{k} < \texttt{max})\ \mathsf{else}\ (\texttt{k} \leq \texttt{min} \lor \texttt{k} \geq \texttt{max}))
			\end{array}
			\right\}
		}
	\end{mathpar}
	\caption{Specifications of \texttt{inRange} for the give-up template}
	\label{fig:inRange_giveup}
\end{figure}

@@

\begin{table}[h]
	\begin{center}
		\begin{tabular}{l | c || l | c } 
			\hline
			\textbf{Lock-coupling template} & \textbf{LoC} & \textbf{Give-up  template} & \textbf{LoC} \\
			\hline
			Template specfications & 190 & Template specfications & 138 \\
			\texttt{findNext} proof 	& 27  & \texttt{findNext} proof & 27 \\
			\texttt{insertOp} proof 	& 32  & \texttt{insertOp} proof & 31 \\
			& & \texttt{inRange} proof & 30 \\
			\texttt{traverse} proof	& 360  & \texttt{traverse} proof & 336 \\ 
			\texttt{insert} proof 	   		& 450  & \texttt{insert} proof & 284 \\
			\texttt{lookup} proof 	   	& 235  &  \texttt{lookup} proof & 304 \\ 
			\texttt{pushdown\_left} proof 	   	& 282 & & \\	
			\texttt{delete} proof 		& 182	 &   & 	 \\
			supporting proofs    & 	1909	 & supporting proofs 	& 1876 \\
			\hline
			\textbf{Total} & \textbf{3667} & & \textbf{3026}  
		\end{tabular}
		\caption{Size of Coq definitions and proofs, by topic.}
		\label{table_coq}
	\end{center}
\end{table}

@@

\section{Proofs for Section \ref{} and \ref{label}}
\label{sec:apd_proof}
Below are partially completed proofs of \texttt{traverse} for the give-up template (Figure \ref{proof_giveup_traverse}), as well as proofs of \texttt{insert} for the lock-coupling and give-up templates, annotated with separation logic specifications. The complete proofs can be found in our Coq mechanization, which is mentioned in Section \ref{proof_mech} for reference.

\begin{figure}[!ht]

@@

		(\mathit{res} = \mathsf{true} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->t} = \texttt{NULL}  \ast \infp (\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \texttt{k} \in \mathsf{range}(\texttt{p'}) \ast \cdots)\end{array}\right\}$

@@

\begin{figure}[!ht]
	$\color{blue}
	m.\left\langle 
	\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m
	\right\rangle$
	\begin{lstlisting}[language = C, style=myStyle, mathescape=true]
void insert (node_t **r, int k, void *v){
		struct pn *pn = (struct pn *) surely_malloc (sizeof *pn);
		pn->n = *r;
		$\color{blue} \left\{\begin{array}{l} \texttt{pn} \mapsto (\texttt{p, r}) \ast  \texttt{r->lock} \lockvar \mathsf{R}(\texttt{r}) \ast \treerep\ m \end{array}\right\}$
		acquire(pn->n->lock);
		$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p, r}) \ast  \texttt{r->lock} \lockvar \mathsf{R}(\texttt{r}) \ast \mathsf{R}(\texttt{r}) \ast \cdots \end{array}\right\}$
		if (traverse(pn, k) == 0){
		$\color{blue} \left\{\begin{array}{l}  \mathit{res} = \mathsf{false} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->t} \neq \texttt{NULL} \ast \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \texttt{k} \in \mathsf{range}(\texttt{p'}) \ast \cdots \end{array}\right\}$
			pn->p->t->value = v;
			$\color{blue} \left\{\begin{array}{l}  \mathit{res} = \mathsf{false} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \texttt{p'->t->value} \mapsto \texttt{v} \ast \cdots \end{array}\right\} \Rrightarrow{\textbf{commit}}$ 
			//Linearization point
			$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \cdots \end{array}\right\}$
		}
		else{
			$\color{blue} \left\{\begin{array}{l}  \mathit{res} =\mathsf{true} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->t} = \texttt{NULL}  \ast \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \texttt{k} \in \mathsf{range}(\texttt{p'}) \ast \cdots \end{array}\right\}$
			insertOp(pn, k, v);
			$\color{blue} \left\{\begin{array}{l}  \mathit{res} =\mathsf{true} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast  \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \\ \texttt{p'->t}\mapsto (\texttt{k, v', l, r})  \ast \texttt{l} \mapsto \texttt{NULL} \ast  \texttt{r} \mapsto \texttt{NULL} \ast \cdots \end{array}\right\} \Rrightarrow{\textbf{commit}}$ 
			//Linearization point
			$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \cdots \end{array}\right\}$
		}
		release(pn->n->lock);
		$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \cdots \end{array}\right\}$
		free(pn);
		$\color{blue} \left\{\begin{array}{l}  \texttt{p'->lock} \lockvar \mathsf{R}(\texttt{p'}) \ast \cdots \end{array}\right\}$
} \end{lstlisting}
	$\color{blue}
	\left\langle 
	\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \texttt{v}])
	\right\rangle$
	\caption{The \texttt{insert} function in lock-coupling template annotated with separation logic specification}
	\label{proof_lock_insert}
\end{figure}

\begin{figure}[!ht]
	$\color{blue}
	m.\left\langle 
	\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ m
	\right\rangle$
	\begin{lstlisting}[language = C, style=myStyle, mathescape=true]
void insert (node_t **r, int k, void *v){
		struct pn *pn = (struct pn *) surely_malloc (sizeof *pn);
		pn->n = *r;
		$\color{blue} \left\{\begin{array}{l} \texttt{pn} \mapsto (\texttt{p, r}) \ast  \infp(\texttt{r}) \ast \treerep\ m \end{array}\right\}$
		if (traverse(pn, k) == 0){
				$\color{blue} \left\{\begin{array}{l}  \mathit{res} = \mathsf{false} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->t} \neq \texttt{NULL}  \ast \infp (\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \texttt{k} \in \mathsf{range}(\texttt{p'}) \ast \cdots \end{array}\right\}$
				pn->p->t->value = v;
				$\color{blue} \left\{\begin{array}{l}  \mathit{res} = \mathsf{false} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \infp (\texttt{p'}) \ast \texttt{p'->t->value} \mapsto \texttt{v} \ast \cdots \end{array}\right\} \Rrightarrow{\textbf{commit}}$ 
				//Linearization point
				$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p', p'}) \ast \infp (\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \cdots \end{array}\right\}$
		}
		else{
				$\color{blue} \left\{\begin{array}{l}  \mathit{res} = \mathsf{true} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast \texttt{p'->t} = \texttt{NULL}  \ast \infp (\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \texttt{k} \in \mathsf{range}(\texttt{p'}) \ast \cdots \end{array}\right\}$
				insertOp(pn, k, v);
				$\color{blue} \left\{\begin{array}{l}  \mathit{res} =\mathsf{true} \ast \texttt{pn} \mapsto (\texttt{p', p'}) \ast  \infp (\texttt{p'}) \ast \\ \texttt{p'->t}\mapsto (\texttt{k, v', l, r})  \ast \texttt{l} \mapsto \texttt{NULL} \ast  \texttt{r} \mapsto \texttt{NULL} \ast \cdots \end{array}\right\} \Rrightarrow{\textbf{commit}}$ 
				//Linearization point
				$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p', p'}) \ast \infp (\texttt{p'}) \ast \mathsf{R}(\texttt{p'}) \ast \cdots \end{array}\right\}$
		}
		release(pn->n->lock);
		$\color{blue} \left\{\begin{array}{l}  \texttt{pn} \mapsto (\texttt{p', p'}) \ast \infp (\texttt{p'}) \ast \cdots \end{array}\right\}$
		free(pn);
		$\color{blue} \left\{\begin{array}{l} \infp (\texttt{p'}) \ast \cdots \end{array}\right\}$
	} \end{lstlisting}
	$\color{blue}
	\left\langle 
	\nodeboxrep\ \texttt{p} \ \big | \ \treerep\ (m[\texttt{k} \mapsto \texttt{v}])
	\right\rangle$
	\caption{The \texttt{insert} function in give-up template annotated with separation logic specification}
	\label{proof_giveup_insert}
\end{figure}
